### 2019-12-17 title_page.Rmd
mansucript adaption to ecography guidelines

---
title: "Sampbias, a method to evaluate geographic sampling bias in species distribution data"
# title: "A global Red List of orchids supports the value of automated conservation assessment"
author: "Alexander Zizka^1,2,3^, Alexandre Antonelli^3,4,5^, Daniele Silvestro^3,4^"
output: 
  bookdown::pdf_document2:
  #word_document:
    toc: false
    toc_depth: 1
    number_sections: false
header-includes: |
    \usepackage{fancyhdr}
    \pagestyle{fancy}
    \fancyhead[L]{sampbias}
    \fancyhead[R]{ }
    \pagenumbering{gobble}
linestretch: 1.3
fontsize: 12pt
---

<!-- Letter to conservation letters: 3000 words in length (excluding references, table and figure legends, and online Supporting Information), 40 references, 6 figures -->

<!-- Contributed paper to Conservation Biology, 3000-6000 words -->

<!-- Affiliations -->
1. German Center for Integrative Biodiversity Research, University of Leipzig, Leipzig, Germany

2. Naturalis Biodiversity Center, Leiden University, Leiden, The Netherlands

3. Gothenburg Global Biodiversity Centre, University of Gothenburg, Gothenburg, Sweden

4. Department of Biological and Environmental Sciences, University of Gothenburg, Gothenburg, Sweden

5. Royal Botanical Gardens Kew, Richmond, Surrey, United Kingdom

**Running head** -- sampbias

**Word count** -- Abstract: XX; Total: XX

<!-- **Article type** -- Letter, 146 words in the abstract, 5400 words in the manuscript, 42 References, 2 tables, 4 Figures -->

**Corresponding author** -- alexander.zizka@idiv.de

<!-- **Authors email addresses** -- tiffany.knight@idiv.de, PVitt@chicagobotanic.org -->

**Keywords** Collection effort, Global biodiversity Information Facility (GBIF), Presence only data, Roadside bias, Sampling intensity

# Acknowledgements
We thank the organizers of the 2016 Ebben Nielsen challenge for inspiring and recognizing this research. We thank all data collectors and contributors to GBIF for their effort.