### 2019-12-18 manuscript.rmd
release version

---
title: "Sampbias, a method to evaluate geographic sampling bias in species distribution data"
#author: "Alexander Zizka^1,2,3^, Alexandre Antonelli^3,4,5^, Daniele Silvestro^3,4^"
bibliography: sampbias.bib
csl: ecography.csl
output: 
  bookdown::pdf_document2:
  #word_document:
    toc: false
    toc_depth: 1
    fig_caption: true
    number_sections: false
header-includes: |
    \usepackage{caption}
    \usepackage{fancyhdr}
    \pagestyle{fancy}
    \fancyhead[L]{sampbias accessibility bias}
    \fancyhead[R]{\thepage}
    \fancyfoot{}
    \usepackage{float}
    \floatplacement{figure}{H}
    \usepackage{lineno}
    \linenumbers
    \usepackage[url=false]{biblatex}
    \usepackage{setspace}
linestretch: 2
fontsize: 12pt
always_allow_html: yes
---

<!-- 1. German Center for Integrative Biodiversity Research, University of Leipzig, Leipzig, Germany -->
<!-- 2. Naturalis Biodiversity Center, Leiden University, Leiden, The Netherlands  -->
<!-- 3. Gothenburg Global Biodiversity Centre, University of Gothenburg, Gothenburg, Sweden -->
<!-- 4. Department of Biological and Environmental Sciences, University of Gothenburg, Gothenburg, Sweden -->
<!-- 5. Royal Botanical Gardens Kew, Richmond, Surrey, United Kingdom -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sampbias)
```

<!-- For Ecography SN, max 4000 (-6000?) words, max 2 figures -->

\newpage{}

# Abstract
Georeferenced species occurrences from public databases have become essential to biodiversity research and conservation, but have limitations. Geographically biased sampling is a widely recognized issue that might severely affect analyses. Especially "roadside bias", i.e. differences in sampling intensity among localities caused by differences in accessibility for humans is ubiquitous and might differ in strength among taxonomic groups and datasets. Yet, no general methodology exists to quantify the effect of roadside or other sources of bias on a dataset level. Here we present *sampbias*, an algorithm and software to estimate the biasing effect of accessibility (by roads, rivers, airports, cities, or any user-defined structures) in species occurrence datasets. *Sampbias* uses a Bayesian approach based on MCMC to optimize the rate of a Poisson sampling process, with sampling dependent on the distance from the next biasing structure. The results are comparable among biasing factors and datasets. *Sampbias* is implemented as a user-friendly R package. We exemplify the use of *sampbias* on a dataset of mammal occurrences from the Indonesian island of Borneo, downloaded from www.gbif.org, showing a high biasing effect of cities and a moderate effect of roads.

# Keywords
Collection effort, Global biodiversity Information Facility (GBIF), Presence only data, Roadside bias, Sampling intensity

\newpage{}

# Background
Publicly available datasets of geo-referenced species occurrences, such as provided by the Global Biodiversity Information Facility (www.gbif.org) have become a fundamental resource in biological sciences, especially in biogeography, conservation, and macroecology. However, these datasets are are typically not collected systematically and rarely include information on collection effort. Instead, they  are often compiled from a variety of sources (e.g. scientific expeditions, census counts, genetic barcoding studies, and citizen-science observations),  therefore often subject to sampling bias [@Meyer2016].

Likely, the number of occurrence available in such datasets is biased by factors other than species' presence or abundance, including the under-sampling of specific taxa ("taxonomic bias", e.g., birds *vs.* nematodes), specific geographic regions ("geographic bias", i.e. easily accessible *vs.* remote areas), and specific temporal periods ["temporal bias", i.e. wet season *vs.* dry season, @Isaac2015; @Boakes2010]. Geographic sampling bias---the fact that sampling effort is spatially biased, rather than equally distributed over the study area---is prevalent in all non-systematically collected datasets of species distributions. Many factors can can cause sampling bias, including socio-economic factors [i.e. national research spending, history of scientific research; @Meyer2015; @Daru2018], political factors [armed conflict, democratic rights; @Ryden2019], and physical accessibility [i.e. distance to a road or river, terrain conditions, slope; @Yang2014; @Botts2011]. Especially physical accessibility is omnipresent as a biasing factor [e.g. @Lin2015; @Engemann2015; @Kadmon2004], across spatial scales, and the term "roadside bias" has been coined for it. In practice, this means that most species observations  are made in or near cities, along roads and rivers, and near other human settlements. Less observations come from the middle of a tropical rainforest or from a mountain top. Interestingly, since the observation of different taxonomic groups has different challenges, geographic sampling bias and the effect of accessibility may differ among taxonomic groups [@Vale2012].

The implications of not considering geographic sampling bias in biodiversity research are likely  substantial [@Meyer2016; @Rocchini2015; @Shimadzu2015;  @Yang2013; @Kramer-Schadt2013; @Barbosa2013]. While the presence of geographic sampling bias is broadly recognized [e.g. @Kadmon2004], and approaches exist to account for them in some analyses---for instance for species-richness estimates [@Engemann2015] species distribution models [@Stolar2015;@Beck2014; @Fithian2014; @Warren2014; @Boria2014; @Varela2014; @Fourcade2014; @Monsarrat2019], occupancy models [@Kery2016], or abundance estimates [@Shimadzu2015]---few attempts have been made to explicitly quantify the bias [@Kadmon2004; @Hijmans2000] or to discern among different sources of bias [@Fernandez2015; @Ruete2015; @Fithian2014], and to our knowledge, no tools exist for comparing the strength of accessibility bias among sources of bias or datasets. While it is unrealistic to expect that spatial biases in biodiversity data will ever disappear, it is crucial that researchers realise the intrinsic bias associated with the  data they are dealing with. This is the first step towards estimating to which extent these biases may affect their analyses, results, and conclusions drawn from such data. Therefore, it is advisable for any study dealing with species occurrence data to assess the strength of accessibility bias in the underlying data.

Here, we present *sampbias*, a method to quantify accessibility bias in individual datasets of species occurrences, in a way that is comparable across datasets. *Sampbias* is implemented as user-friendly R-package. Specifically, *sampbias* uses a Bayesian approach based on a Poisson process to address three questions:

1) How strong is the accessibility bias in a given dataset?

2) How important are different means of human accessibility, such as to airports, cities, rivers or roads, in causing this bias?

3) How is sampling bias distributed in space, i.e. which areas are a priority for targeted sampling?

*Sampbias* is implemented in R [@rcoreteam2019], based on commonly used packages for data handling  [`ggplot`, @Wickham2009; `tidyr`, @Wickham2019a; `dplyr`, @Wickham2019; `forcats`, @Wickham2019b; `magrittr`, @Bache2014; `viridis`, @Garnier2018], handling geographic information and geo-computation [`raster`, @Hijmans2019; `sp`, @Pebesma2005; @Bivand2013] and statistical modelling [`stats`, @rcoreteam2019]. *Sampbias* offers an easy and largely automated means for biodiversity scientists and non-specialists alike to explore bias in species occurrence data and may be used to identify priorities for further collection or digitalization efforts, provide bias surfaces for species distribution modelling, or assess the reliability of scientific results based on publicly available species distribution data.

# Methods and Features
## General concept
Under the assumption that organisms exist across the entire area of interest, we can expect the number of sampled occurrences in a restricted areas, such as a single biome, to be distributed uniformly in space (even though, of course, the density of individuals and the species composition may be heterogeneous). With *sampbias* we assess if discrepancies in sampling can be explained by distance from factors that potentially bias their sampling probability (e.g. cities or roads).

*Sampbias* works on a user-defined scale, and any dataset of multi-species occurrence records can be tested against any geographic gazetteer (reliability increases with increasing dataset size). Default large-scale gazetteers for airports, cities, rivers and roads are provided with *sampbias*. Species occurrence data as downloaded from the data portal of GBIF can be directly used as input data for sampbias. The output of the package includes measures of bias effect, which are comparable between different gazetteers (e.g. comparing the biasing effect of roads and rivers), different taxa (e.g. birds *vs.* flowering plants) and different data sets (e.g. specimens *vs.* human observations).

## Distance calculation
*Sampbias* uses gazetteers of the geographic location of bias sources (e.g. roads) to generate a regular grid across the study area (the geographic extent of the dataset). For each grid cell $i$, we then compute a vector $X_i(j)$ of minimum distances ("as the crow flies") to each source of bias $j \in B$. We then use these distance grids to sample the distribution of distances in the observed dataset. The resolution of the grid defined the precision of the distance estimates, for instance a 1x1 degree raster will yield approximately a 100km precision at the equator.  Due to the assumption of homogeneous sampling and a computational trade-off between the resolution of the distance raster and the extent of the study area (for instance, a 1000m resolution for a global dataset would lead to the generation of grid  for which distance calculation will become computationally prohibitive in most practical cases), *sampbias* is best suited for local or regional datasets at high resolution (c. 100 -- 10,000m).

## Quantifying accessibility bias using a Bayesian framework
We describe the observed number of sampled occurrences $S_i$ within each cell $i$ as the result of a Poisson sampling process with rate $\lambda_i$. We model the rate $\lambda_i$ as a function of a constant $q$, which represent the expected number of occurrences per cell in the absence of biases, i.e. when $\sum_{j=1}^{B}{ X_i(j)} = 0$. Additionally, we model $\lambda_i$ to decrease exponentially as a function of distance from sources of bias, such that increasing distances will result in a lower sampling rate. For a single source of bias the rates of cell $i$ with distance $X_i$ from a bias is:
$$
\lambda_i = q \times \exp \left(- w X_i \right)
$$
where $w \in \mathbb{R^+}$ defines the steepness of the Poisson rate decline, such that $w \approx 0$ results in a null model of uniform sampling rate $q$ across cells. In the presence of multiple bias predictors, the sampling rate decrease is a function of the cumulative effects of each bias and its distance from the cell:
$$
\lambda_i = q \times \exp \left(-\sum_{j=1}^{B}{w_j X_i(j)} \right) \ \ \ \ (1)
$$
where a vector $\mathbf{w} = [w_1, ..., w_B]$ describes the amount of bias attributed to each specific predictor.

To quantify the amount of bias associated with each predictor, we jointly estimate the parameters $q$ and $\mathbf{w}$ in a Bayesian framework. We use Markov Chain Monte Carlo (MCMC) to sample these parameters from their posterior distribution:
$$
P(q, \mathbf{w} | \mathbf{S}) \propto \prod_{i=1}^{N}{ Poi(S_i | \lambda_i) } \times P(q) P(\mathbf{w}) \ \ \ \ (2)
$$
where the likelihood of sampled occurrences $S_i$ within each cell $Poi(S_i | \lambda_i)$ is the probability mass function of a Poisson distribution with rate per cell defined as in Eqn. (1). The likelihood is then multiplied across the $N$ cells considered. We used exponential priors on the parameters $q$ and $\mathbf{w}$, $P(q) \sim \Gamma(1, 0.01)$ and $P(\mathbf{w}) \sim \Gamma(1, 1)$, respectively.

We summarize the parameters by computing the mean of the posterior samples and their standard deviation. We interpret the magnitude of the elements in $\mathbf{w}$ as a function of the importance of the individual biases. We note however that this test is not explicitly intended to assess the significance of each bias predictor (for which a Bayesian variable selection could be used), particularly since several sources of bias might be correlated (e.g. cities, and airports). Instead, these analyses can be use to quantify the expected amount of bias in the data that can be predicted by single or multiple predictors in order to identify under-sampled and unexplored areas.

[**Daniele some text here on the projection of bias through space, and what the plots exactly show**]

## Example and Empirical analysis
A default *sampbias* analysis can be run with few lines of code in R. The main function `calculate_bias` creates an object of the class `"sampbias"`, for which the package provides a plotting and summary method. Based on a `data.frame` including species identity and geographic coordinates. Additionally, some options exist to provide custom gazetteers, custom distances for the bias estimation, a custom grain size of the analysis, as well as some operators for the calculation of the bias distances. A tutorial on how to use sampbias is available with the package and in the electronic supplement of this publication (Appendix S1).

To exemplify the use and output of *sampbias*, we downloaded the occurrence records of all mammals available from the Indonesian island of Borneo [n = 6,262, @gbifdoi2016], and ran *sampbias* using the default gazetteers as shown in the example code below, to test the biasing effect of the main airports, cities and roads in the dataset. The example dataset is provided with the *sampbias*. We found a strong effect of cities on sampling intensity, a moderate effect of roads and airports and no effect of rivers (Fig. \ref{fig:empirical}A). All models predict a low number of collection records in the centre of Borneo, which reflects the original data, and where accessibility means are low (Figure S1 in Appendix S1). The empirical example illustrates the use of *sampbias*, for detailed analyses or a smaller geographic scale, higher resolution gazetteers, including smaller roads and rivers and a higher spatial resolution would be desirable. Results might change with increasing resolution, since roads and rivers might have a stronger effect on higher resolutions (facilitating most the access to their immediate vicinity), whereas cities and airports might have a stronger effect on the larger scale (facilitating access to a larger area).

```{r, eval = FALSE}
library(sampbias)

#a data table with species identify, longitude, and latitude
example.in <- read.csv(system.file("extdata", 
                                   "mammals_borneo.csv",
                                   package="sampbias"), 
                       sep = "\t")

#running sampbias
example.out <- calculate_bias(x = example.in, 
                              res = 0.05,
                              buffer = 0.5)

# summarizing the results
summary(example.out)
plot(example.out)

#project in space
proj <- project_bias(example.out)
map_bias(proj)
```

# Data accessibility
*Sampbias* is available under a GPL-3 license from https://github.com/azizka/sampbias, and includes an example dataset as well as a tutorial to run *sampbias* (Appendix S2) and as summary of possibly warnings produced by the package (Appendix S3).

<!-- # Acknowledgements -->
<!-- We thank the organizers of the 2016 Ebben Nielsen challenge for inspiring and recognizing this research. We thank all data collectors and contributors to GBIF for their effort. -->

# Author contributions
All authors conceived of this study, AZ and DS developed the statistical algorithm, AZ and DS wrote the R-package, AZ and DS wrote the manuscript with contributions from AA.

# Figure
<!-- empirical results -->
```{r empirical, out.width = "6in", echo = FALSE, fig.cap = "The spatial projection of the accessibility bias in an empirical example dataset of mammal occurrences on the Indonesian island of Borneo from www.gbif.org. A) bias weights, B) projection of the expected number of occurrences given the sampbias model. At the study scale of 0.05 degrees (~5 km)\\textit{Sampbias} finds the strongest biasing effect for the proximity of cities and roads, and a highest undersampling in the center of the island.", eval = TRUE}
knitr::include_graphics(path = "ms_figures/figure_empirical_results.jpg")
```

# Supplementary material

Appendix S1 - Supplementary Figure

Appendix S2 - Tutorial running sampbias in R

Appendix S3 - Possible warnings and their solutions

\newpage{}
# References