### 2020-09-15 04_appendix_S1.Rmd
tidy up github

---
title: "sampbias, a method for quantifying geographic sampling biases in species distribution data"
subtitle: "Appendix S1 - Supplementary Figures"
#author: "Alexander Zizka, Alexandre Antonelli, Daniele Silvestro"
output: rmarkdown::pdf_document
header-includes: |
    \usepackage{float}
    \renewcommand{\thefigure}{S\arabic{figure}}
---

\newpage{}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
```

```{r, fig.height = 8, fig.cap="The example dataset of mammal occurrences from the island of Borneo, as downloaded from www.gbif.org (n = 6,262, 2016, https://doi.org/10.15468/dl.7fg4zx), and the geographic gazetters of main cities, roads, rivers and airports used for the sampbias analysis. Note that the records in the sea around Borneo were excluded by with the help of the user-provided study area."}
#libraries
library(raster)
library(ggplot2)
library(magrittr)
library(sp)
library(dplyr)
library(viridis)
library(readr)
library(sf)
library(geosphere)
library(tidyr)
library(forcats)

## get a polygon of Borneo
data(borneo)

#load data
occ <-read.csv(system.file("extdata", "mammals_borneo.csv", package="sampbias"), sep = "\t") %>% 
  dplyr::select(latitude = decimalLatitude, longitude = decimalLongitude) %>% 
  mutate(type = "Mammal occurrence")

data(airports)
data(waterbodies)
data(cities)
data(roads)
data(landmass)

# prepare for plotting
airp <- airports %>% crop(SpatialPoints(occ[,2:1])) %>% raster::intersect(borneo) %>%  coordinates() %>%  data.frame() %>% mutate(type = "Airport")
cit <- cities %>% crop(SpatialPoints(occ[,2:1])) %>% raster::intersect(borneo) %>%  coordinates() %>%  data.frame() %>%  mutate(type = "City")
pts <- bind_rows(occ, cit, airp)

riv <- waterbodies %>% crop(SpatialPoints(occ[,2:1])) %>%  fortify() %>%  mutate(type = "River")
road <- roads %>% crop(SpatialPoints(occ[,2:1])) %>% raster::intersect(borneo) %>%  fortify() %>%  mutate(type = "Road")
lin <- bind_rows(riv, road)

lma <- borneo %>%  fortify()

# plot
ggplot()+
  geom_polygon(data = lma,
               mapping = aes(x = long, y = lat, group = group), lwd = 0.5, col = "grey40", fill = "transparent")+
  geom_path(mapping = aes(x = long, y = .data$lat, group = .data$group, shape = type, color = type),
            data = lin)+
  geom_point(data = pts, mapping = aes(x = longitude, y = latitude, linetype = type, color = type), size = 3)+
  scale_color_viridis(discrete = TRUE)+
  xlab("Longitude")+
  ylab("Latitude")+
  coord_fixed()+
  theme_bw()+
  theme(
    legend.position = "bottom"
  )
```

```{r}
occ <-read.csv(system.file("extdata", "mammals_borneo.csv", package="sampbias"), sep = "\t")
area <- round(areaPolygon(borneo) / 1000000, 0)
```

```{r, echo=FALSE}

inp <- paste("empirical_analyses/simulations/", list.files("empirical_analyses/simulations/", pattern = "weightestimate"), sep = "")
inp <- lapply(inp, "read_csv")

dat <- bind_rows(inp) %>% 
  mutate(records_area = round(rar / area * 10000, 2)) %>% 
  dplyr::select(contains("w"), ID, type, records_area) %>%
  pivot_longer(cols = contains("w_"), names_to = "bias",
                 values_to = "posterior_estimate") %>%
    mutate(bias = gsub("w_", "", .data$bias)) %>% 
    mutate(bias = fct_reorder(.data$bias, .data$posterior_estimate, .fun = median, .desc = FALSE))

```

```{r, echo = FALSE, fig.height = 7, fig.cap="The bias weights ($w$) of the empirical analyses of records from mammals from Borneo (6262 records, www.gbif.org, 2016, https://doi.org/10.15468/dl.7fg4zx) in comparison to ten simulated data sets of the same size with randomly picked locations across Borneo (no bias)."}

emp <- dat %>% 
  filter(records_area == 85) %>% 
  mutate(ID = "Empirical") %>% 
  mutate(type = "Empirical") %>%
  dplyr::select(-records_area)

sim <- paste("empirical_analyses/simulations/", list.files("empirical_analyses/simulations/", pattern = "biasweights"), sep = "")
sim <- lapply(sim, "read_csv")

plo <- bind_rows(sim) %>%
  mutate(ID = paste("Rep_", as.character(ID), sep = "")) %>% 
  mutate(type = "Simulated (no bias)") %>% 
  dplyr::select(contains("w"), ID, type) %>%
  pivot_longer(cols = contains("w_"), names_to = "bias",
                 values_to = "posterior_estimate") %>%
    mutate(bias = gsub("w_", "", .data$bias)) %>% 
  bind_rows(emp)

ggplot(data = plo)+
  geom_boxplot(aes(x = ID, 
                   y = posterior_estimate,
                 fill = as.factor(type),
                 color = as.factor(type)))+
  ylab("Posterior weight")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw()+
  theme(legend.title = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "bottom")+
  facet_wrap(.~bias, ncol = 1)
```

```{r, echo = FALSE, fig.cap="The bias weights ($w$) estimated from data sets with differing density of occurrence records across the study area. Datasets generated by down-sampling the empirical example dataset of mammal occurrences on Borneo (www.gbif.org,  2016, https://doi.org/10.15468/dl.7fg4zx)."}
ggplot(data = dat)+
  geom_boxplot(aes(x = bias, 
                   y = posterior_estimate,
                 fill = as.factor(records_area),
                 color = as.factor(records_area)))+
  ylab("Posterior weight")+
  xlab("Bias factor")+
  scale_fill_viridis(discrete = TRUE, name = "Records per\n10,000 sqkm")+
  scale_color_viridis(discrete = TRUE, name = "Records per\n10,000 sqkm")+
  theme_bw()
```

```{r, out.width = "6in", fig.cap = "Projected bias surface for the distribution records of mammals from Borneo (www.gbif.org, 2016, https://doi.org/10.15468/dl.7fg4zx). A) - D) Different levels of random downsampling. The plotted extent is adapted to the study area, which may not cover the entire island for rarified datasets with few records (A) and B))"}

knitr::include_graphics(path = "ms_figures/figure_projection_empirical_rarefaction.jpg")

```